---
title: "lab3"
author: "Clara Chun"
date: "June 8, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Importing the Data
Run install the package `readr` by running `install.packages("readr")` in the console.

The code below reads in the `tents.csv` comma-separated file.  The `col_types` argument is telling `R` what the data type is in each column.  See `?read_csv` if you're curious.
```{r}
library(readr)
tents <- read_csv("tents.csv", col_types = "ccnniccc")
```

### Questions
Answer the following questions programatically.  Some questions ask for you to explain what you see---you may answer this as text outside a code block.

1. Run `str()` on `tents`.  What information does it tell you? This is a useful function to know.
```{r}
str(tents)
```
The str() function shows the first few values of each of the variables. It lists the information about the data set.

2. What class is `tents`?
```{r}
class(tents)
```


3. How many rows are in `tents`?
```{r}
nrow(tents)
```


4. How many columns are in `tents`?
```{r}
ncol(tents)
```


5. What are the names of the columns?
```{r}
colnames(tents)
```


6. Display the first 5 rows.
```{r}
head(tents, 5)
```


7. Display the last 4 rows.
```{r}
tail(tents, 4)
```

8. Find out how to use `quantile` to find the interquartile range (75th percentile - 50th percentile) of the weights of the tents.
```{r}
quantile(tents$weight, probs = 0.75) - quantile(tents$weight, probs = 0.25)
```

9. Find the standard deviation of the heights of the `tents`.  Does `R` use `n` or `n-1` in the denominator when calculating standard deviation?
```{r}
sd(tents$height)
```
`R` uses `n-1` in the denominator when calculating standard deviation. 

10. The weight is given in grams.  Add a new column to the data frame called `weight_lbs` that contains the weight of the tents in pounds.  Remember to overwrite `tents` with your update.
```{r}
tents$weight_lbs = 0.0625*tents$weight
tents
```

11. The height is given in centimeters.  Add a new variable `height_in` to the data frame that contains the height of the tents in inches. Remember to overwrite `tents` with your update.
```{r}
tents$height_in = 0.393701*tents$height
tents
```


12. How many tents cost more than $400?
```{r}
length(tents$price > 400)
```

13. Subset the data frame so that it only contains tents that are valued $300 or less.  Assign the subsetted data frame the name `so_cheap_tents`.
```{r}
so_cheap_tents = tents[tents$price < 300,]
so_cheap_tents

```


14. Of the tents that are more than $300, what percent are `big-agnes` tents?
```{r}
sum(tents$price > 300 & tents$brand == "big-agnes")/sum(tents$price > 300)

```


15. Select the `bestuse` column of `tents` in four different ways: using `$`, `[[]]`, position number and `[]`, and variable name and `[]`. 
```{r}
tents$bestuse
tents[["bestuse"]]
tents["bestuse"]
```


16. What is the name of the tent with the best height-to-weight ratio?
```{r}
tents$name[which.max(tents$height/tents$weight)]
```


17. Run `table` with two arguments, `seasons` and `capacity` (this should be one table).  What did it do?
```{r}
table(tents$seasons, tents$capacity)
```
It created a table of the counts of the two varibles `seasons` and `capacity`

18. What type of tent (bestuse) is the most common? Your answer should return only the use and not the count.
```{r}
which.max(table(tents$bestuse))
```


19. Create a data frame called `kelty` with only data on kelty tents.  Do the same for `big-agnes`.  Then, combine the two data frames.
```{r}
kelty_frame = tents[tents$brand == "kelty",]
bigagnes_frame = tents[tents$brand == "big-agnes",]
rbind(kelty_frame, bigagnes_frame)
```


