---
title: "lab6"
author: "Clara Chun"
date: "June 15, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Population Variance

Write a function `MyVar` that calculates the population variance of a vector.  If you need a refresher of what that means, [this section of wikipedia should help](https://en.wikipedia.org/wiki/Variance#Population_variance_and_sample_variance).

```{r}
MyVar = function(x) {
  mu = sum(x)/length(x)
  ((1/length(x)) * sum(x^2)) - mu^2
}

```
Test you function:
```{r}
MyVar(1:3)
```


## Finding Roots
Write a function `zeroes` that finds the roots (zeroes) of a degree two polynomial.  The function should take the coefficients of the polynomial $a^2x + bx + c$ as its arguments and returns a vector of roots.  If there are no real roots, the function should stop and produce an error message: `"This polynomial has no real roots."`

```{r}
zeroes = function(a, b, c) {
  if ((b^2-4*a*c)<0){
    print("This polynomial has no real roots")
  } else{
    root1 = (-b + sqrt(b^2 - 4*a*c))/2*a
    root2 = (-b - sqrt(b^2 - 4*a*c))/2*a
    return(c(root1, root2))
  }
}
```

Test your function:
```{r}
zeroes(1, 0, 0)
zeroes(1, 0, 1)
zeroes(1, 1, 1)
```



## Temperature Conversion

Write a function `ConvertTemperature` that takes a numeric vector of temperatures in Fahrenheit and converts them into Celsius, Kelvin, Rankine, Delisle, Newton, Reamur, or Romer, depending on what the user specifies in an argument called `to`.  If they specify some other temperature scale, return the message "The [their input] temperature scale is not supported".  If you don't know the conversions, (see this table)[https://en.wikipedia.org/wiki/Conversion_of_units_of_temperature#Fahrenheit]

```{r}
ConvertTemperature = function(x, to){
  if (to == "Celsius") {
    (x - 32)*(5/9)} 
  else if (to == "Kelvin") {
    (x + 459.67)*(5/9)} 
  else if (to == "Rankine") {
    x + 459.67} 
  else if (to == "Delisle") {
    (212 - x)*(5/6)} 
  else if (to == "Newton") {
    (x - 32)* (11/60)} 
  else if (to == "Reamur") {
    (x - 32)*(4/9)} 
  else if (to == "Romer") {
    (x - 32)*(7/24) + 7.5} 
  else {
    stop (str_c(to, "is not supported", sep = " "))}
}

```

Test your function
```{r}
ConvertTemperature(32, "Celsius") == 0
ConvertTemperature(212, "Kelvin") == 373.15
ConvertTemperature(10, "Rankine") == 501.67
```


## Rolling Dice

For this problem, you'll need the `sample` and `pmax` function.  `sample` generates a random sample of a specified size either with or without replacement.  You can specify probability weights for each element of your sample space using the `prob` argument.  By default, each element will have equal probability.
```
sample(x = vector_to_be_sampled, size, replace = FALSE, prob = NULL)
```
1. What does `pmax` do? How is it different from `max`? Hint: test both on the two vectors `c(1, 5, 3)` and `c(2, 4, 2)`. __`pmax` returns the parallel maxima of the given vector whereas max returns one maximum value of the given vector__
```{r}
pmax(c(1, 5, 3))
pmax(c(2, 4, 2))
max(c(1, 5, 3))
max(c(2, 4, 2))
```

2. When you "roll dice with advantage", you roll a pair of dice and take the higher value as your roll.  Write a function `AdvRoll` the simulates rolling with advantage `n` times. The function should take on 2 arguments `sides` and `n` and output a length `n` vector of rolls.
```{r}

AdvRoll= function(sides,n) {
      
replicate(n,expr=max(sample(c(1:sides),2)))

}
```
Test you function:
```{r}
AdvRoll(6,1000)
```


3. Write your own summary function called `MySummary` that, given a numeric vector `x`, returns the following summary statistics: mean, median, mode, sum, and sum of squares.
```{r}
MySummary = function(x) {
  mymode = function(x) {
    uniqm = unique(x)
    uniqm[which.max(tabulate(match(x,uniqm)))]
  }
  c(mean = mean(x), 
  median = median(x), 
  sum = sum(x),
  mode = mymode(x),
  SumOfSquares = sum(x^2))
}
```
Test your function:
```{r}
MySummary(c(1,1,2,3))
```

4. Using your two functions, simulate "rolling dice with advantage" 1000 times and find the summary statistics.

```{r}
MySummary(AdvRoll(6,1000))
```


