---
title: "lab25"
date: "July 26, 2016"
output: html_document
---

Write a function called `moveCars()` that moves the system one full step; that is, the function should move the red cars east once and the blue cars north once, with red cars going first.

Your function should take the following inputs:
`m`: a matrix

Your function should return a list containing two elements:
`m`: an updated matrix
`not_gridlocked`: a logical value that should be TRUE if the system changed and FALSE if not

```{r}
moveCars = function(m) {
  
  redCars = m * (m ==1)
  blueCars = m * (m ==2)
  
  red_test = m[, c(2:nrow(m),1)]
  
  notblocked_red = m * (red_test == 0)
  notblocked_red = notblocked_red * (notblocked_red == 1)
  
  blocked_red = redCars * (m[, c(2:nrow(m),1)] != 0)
  
  red_move = notblocked_red[,c(nrow(m),c(1:nrow(m)-1))] + blocked_red + blueCars
  
  
  n_test <- red_move[c(nrow(m),c(1:nrow(m)-1)),]
  
  blocked_blue = red_move * (red_move == 2) * (n_test != 0)
  notblocked_blue = red_move * (red_move ==2) * (n_test == 0)
  
  if(nrow(m) == 1)
    m_new = notblocked_blue[1,] + blocked_blue + red_move*(red_move == 1)
  else
    m_new = notblocked_blue[c(2:nrow(m),1),] + blocked_blue + red_move*(red_move == 1)
  
  return(list(m_new, !all(m == m_new)))

}
```


Check:
```{r, eval = FALSE}
m <- matrix(c(1,0,0,0,2,1,2,0,0,2,0,1,1,0,1,2,0,1,2,1,0,0,1,1,2), nrow = 5, ncol = 5, byrow = T)
identical(moveCars(m), list(matrix(c(0,1,0,0,2,1,2,0,0,2,1,1,0,1,0,2,0,1,2,1,0,0,1,1,2), nrow = 5, ncol = 5, byrow = T), TRUE))

m <-  matrix(c(0,1,0,2,1), nrow = 1, ncol = 5, byrow = T)
identical(moveCars(m), list(matrix(c(1,0,1,2,0), nrow = 1, ncol = 5, byrow = T), TRUE))

m <- matrix(c(2,0,1,2,0), nrow = 5, ncol = 1, byrow = T)
identical(moveCars(m), list(matrix(c(0,0,1,2,2), nrow = 5, ncol = 1, byrow = T), TRUE))
```

Now write a function called `initGrid()` that randomizes the initial system.

Your function should take the following inputs:
`r`: number of rows in the matrix
`c`: number of columns in the matrix
`d`: density of the grid 

Your function should return the following:
`m`: an rxc matrix with entries 0 (no cars), 1 (red cars), or 2 (blue cars)

```{r}
initGrid <- function(r,c,d) {
  
  m <- matrix(sample(c(0,1,2), prob = c(1-d,d/2,d/2)),r,c)
  return(m)
} 
```






